---
title: "9041 Assignment 2"
author: Yiyao Wang
date: 2026-10-01
format: 
    html:
      toc: true
      code-fold: true
      code-summary: "Show code"
      df-print: paged
execute:
    echo: false
    warning: false
---

```{r, load libraries}
#| output: FALSE
library(tidyverse)
library(readxl)
library(psych)
library(gt)
library(janitor)
```

```{r, load data}
#| output: FALSE
# loading data
data<-read_excel("./Attachment_Anxiety_Data.xlsx")

# inspecting data
head(data)
```

```{r, clean data}
#| output: FALSE

# clean data 
data.clean <- data |>
  mutate(Gender = case_when(
    Gender == 1 ~ "Male",
    Gender == 2 ~ "Female",
    TRUE ~ NA
  )) |>
  mutate(`Age group` = case_when(
    `Age group` == 1 ~ "18-24",
    `Age group` == 2 ~ "24-34",
    `Age group` == 3 ~ "35-44",
    `Age group` == 4 ~ "45-54",
    `Age group` == 5 ~ "55-64",
    TRUE ~ NA
  )) |>
  mutate_at(7:45, as.integer)


```

```{r, coding the scales}
#| output: FALSE
SA_key <- list(
  Social_Anxiety = c(
    1, 2, 3, 4, -5, 6, 7, 8, -9, 10, -11,
    12, 13, 14, 15, 16, 17, 18, 19, 20
  )
)

AA_key <- list(
  Attachment_Anxiety = c(
    1, 2, 3, 4, 5, 6, 7, 8, 9
  )
)

SEst_key <- list(
  Self_Esteem = c(
    1, -2, 3, 4, -5, -6, 7, -8, -9, 10
  )
)

# social interaction anxiety scale
scores_SA <- scoreItems(
  SA_key,
  data.clean[, paste0("SA_", 1:20)],
  min = 0,
  max = 4
)

scores_SA.df <- as.data.frame(scores_SA$scores)


# attachment anxiety scale
scores_AA <- scoreItems(
  AA_key,
  data.clean[, paste0("AA_", 1:9)],
  min = 1,
  max = 7
)

scores_AA.df <- as.data.frame(scores_AA$scores)

# rosenberg self-Esteem scale
scores_SEst <- scoreItems(
  SEst_key,
  data.clean[, paste0("SEst_", 1:10)],
  min = 1,
  max = 4
)

scores_SEst.df <- as.data.frame(scores_SEst$scores)

allscores.df <- cbind(
  scores_SA.df,
  scores_AA.df,
  scores_SEst.df
)

combined.df <- cbind(data.clean, allscores.df)

final.df <- combined.df |>
  select(Gender, `Age group`,
         Social_Anxiety,
         Attachment_Anxiety,
         Self_Esteem)

```

```{r, demographic table}
#| output: FALSE

demographic.table <- final.df |>
  tabyl(Gender, `Age group`)

demographic.gt <- demographic.table |>
  gt() |>
  tab_header(
    title = "Participant Demographics"
  )

demographic.gt
```

```{r, alpha value table}
#| output: FALSE

summary <- final.df |>
  select(Social_Anxiety, Attachment_Anxiety, Self_Esteem) |>
  describe()

sample_size <- nrow(final.df)

summary$Scales <- rownames(summary)

alphas <- round(c(
  scores_SA$alpha,
  scores_AA$alpha,
  scores_SEst$alpha
), 2)

summary <- summary |>
  select(Scales, mean, median, sd, range)

summary <- cbind(summary, Alpha = alphas)


score.table <- summary |>
  gt() |>
  tab_header(
    title = "Summary Statistics for Psychological Scales"
  )
score.table

```

```{r, filter my data}
#| output: FALSE
filtered.df <- final.df |>
  filter(Gender == "Male")

# summary statistics for male participants
summary_filtered <- filtered.df |>
  select(Social_Anxiety, Attachment_Anxiety, Self_Esteem) |>
  describe()

# sample size
sample_size_filtered <- nrow(filtered.df)

# add scale names
summary_filtered$Scales <- rownames(summary_filtered)

# keep required columns
summary_filtered <- summary_filtered |>
  select(Scales, mean, median, sd, range)

# use the alpha values calculated previously
alphas_filtered <- alphas

# combine summary statistics and alpha
filtered.reptable <- cbind(summary_filtered, Alpha = alphas_filtered)

# create gt table
filtered.score.table <- filtered.reptable |>
  gt() |>
  tab_header(
    title = "Summary Statistics for Male Participants"
  )

filtered.score.table
```


## Intro
The purpose of this analysis was to examine participants’ social interaction anxiety, attachment anxiety, and self-esteem using psychological survey data. The data were cleaned and scored in R. Descriptive statistics and internal consistency estimates were calculated for each psychological scale, and an additional analysis was conducted for male participants.

## Methods

### Participants
The dataset contained `{r} sample_size` participants. Participant gender and age group were used to describe the demographic composition of the sample. The distribution of participants across these demographic categories is shown below.
`{r} demographic.gt`

### Measures
Social Interaction Anxiety. Social interaction anxiety was measured using the Social Interaction Anxiety Scale. Items 5, 9, and 11 were reverse-keyed prior to calculating participants’ average scale scores. The scale demonstrated an internal consistency of α = `{r} alphas[1]`.

Attachment Anxiety. Attachment anxiety was measured using the Experiences in Close Relationships Anxiety Subscale. Participants’ responses were averaged to produce an attachment anxiety score. The scale demonstrated an internal consistency of α = `{r} alphas[2]`.

Self-Esteem. Self-esteem was measured using the Rosenberg Self-Esteem Scale. Items 2, 5, 6, 8, and 9 were reverse-keyed prior to calculating participants’ average scale scores. The scale demonstrated an internal consistency of α = `{r} alphas[3]`.

### Data Analysis Strategy
Survey responses were cleaned and demographic variables were assigned informative labels prior to analysis. Scale scores were calculated separately for social interaction anxiety, attachment anxiety, and self-esteem using scoreItems(). Reverse-keyed items were specified when scoring the Social Interaction Anxiety Scale and Rosenberg Self-Esteem Scale. Average scores and Cronbach’s alpha values were then calculated for each scale. Descriptive statistics were calculated for the full sample and subsequently for a filtered dataset containing only male participants.

## Results
Descriptive statistics for social interaction anxiety, attachment anxiety, and self-esteem in the full sample are presented below.

`{r} score.table`

Across the full sample, the mean social interaction anxiety score was `{r} round(summary$mean[summary$Scales == "Social_Anxiety"], 2)`, the mean attachment anxiety score was `{r} round(summary$mean[summary$Scales == "Attachment_Anxiety"], 2)`, and the mean self-esteem score was `{r} round(summary$mean[summary$Scales == "Self_Esteem"], 2)`.

To examine a subset of the data, the sample was filtered to include only male participants. The filtered sample contained `{r} sample_size_filtered` participants. Descriptive statistics for this group are presented below.

`{r} filtered.score.table`

Among male participants, the mean social interaction anxiety score was`{r} round(summary_filtered$mean[summary_filtered$Scales == "Social_Anxiety"], 2)`  , the mean attachment anxiety score was `{r} round(summary_filtered$mean[summary_filtered$Scales == "Attachment_Anxiety"], 2)`, and the mean self-esteem score was `{r} round(summary_filtered$mean[summary_filtered$Scales == "Self_Esteem"], 2)`.
